Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
rule permutation

($\cfaD/\fequ$)

($\wrD/\fequ$)

($\fequ/\cfaD$)

($\fequ/\me$)

\begin{equation*}
\vlderivation{
  \vlin{\me}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
    \vlin{\fequ}{(x \notin fv(B \vlan D))}{\forall x.((A \vlan C) \vlor (B \vlan D))}{
      \vlhy{(forall x.(A \vlan C)) \vlor (B \vlan D)}}}}
\end{equation*}

($\fequ/\mfaD$)

($\fequ/\mexD$): similar to $\fequ/\mfaD$

@@

  (see~\cite{str:medial,lamarche:gap} for a general discussion on

@@

>>>>

@@

  (see~\cite{str:07:RTA,lamarche:gap} for a general discussion on

@@

Interaction between two rules:
\begin{itemize}
\item $\cfaD/\fequ$:

\item $\wrD/\fequ$:

\item $\fequ/\cfaD$:

\item $\fequ/\me$:

\begin{equation*}
\vlderivation{
  \vlin{\me}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
    \vlin{\fequ}{(x \notin fv(B \vlan D))}{\forall x.((A \vlan C) \vlor (B \vlan D))}{
      \vlhy{(\forall x.(A \vlan C)) \vlor (B \vlan D)}}}}
\end{equation*}

\item $\fequ/\mfaD$:

\item $\fequ/\mexD$: similar to $\fequ/\mfaD$

Interactions between two non-$\fequ$ rules with the presence of $\fequ$ in
between:

\begin{itemize}
\item $\cfaD/\fequ/\rr$ where $\rr \in \set{\me, \mfaD, \mexD}$: First permute
$\cfaD$ under $\fequ$ and then permute $\cfaD$ under $\rr$.

\item $\acD/\fequ/\rr$ where $\rr \in \set{\me, \mfaD, \mexD}$: First permute
$\acD$ under $\fequ$ and then permute $\acD$ under $\rr$.

\item $\wrD/\fequ/\cfaD$: We only list non-trivial cases here:
  \begin{equation*}
  \vlderivation{
    \vlin{\cfaD}{}{(\forall x.A) \vlor B}{
      \vlin{\fequ}{(x \notin fv(B))}{(\forall x.\forall x.A) \vlor B}{
        \vlin{\wrD}{}{\forall x.((\forall x.A) \vlor B)}{
          \vlhy{\forall x.\forall x.A}}}}}
  \leadsto
  \vlderivation{
    \vlin{\wrD}{}{(\forall x.A) \vlor B}{
      \vlin{\cfaD}{}{(\forall x.A) \vlor B}{
        \vlhy{\forall x.\forall x.A}}}}
  \end{equation*}
  \begin{equation*}
  \vlderivation{
    \vlin{\cfaD}{}{(\forall x.A) \vlor B}{
      \vlin{\fequ}{(x \notin fv(B))}{(\forall x.\forall x.A) \vlor B}{
        \vlin{\wrD}{}{\forall x.(B \vlor (\forall x.A))}{
          \vlhy{\forall x.B}}}}}
  \leadsto
  \vlderivation{
    \vlin{\fequ}{(x \notin fv(B))}{(\forall x.A) \vlor B}{
      \vlin{\wrD}{}{\forall x.(B \vlor A)}{
        \vlhy{\forall x.B}}}}
  \end{equation*}

  \begin{equation*}
  \vlderivation{
    \vlin{\cfaD}{}{\forall x.(A \vlor B)}{
      \vlin{\fequ}{(x \notin fv(B))}{\forall x.\forall x.(A \vlor B)}{
        \vlin{\wrD}{}{\forall x.((\forall x.A) \vlor B)}{
          \vlhy{\forall x.\forall x.A}}}}}
  \leadsto
  \vlderivation{
    \vlin{\wrD}{}{\forall x.(A \vlor B)}{
      \vlin{\cfaD}{}{\forall x.A}{
        \vlhy{\forall x.\forall x.A}}}}
  \end{equation*}

  \begin{equation*}
  \vlderivation{
    \vlin{\cfaD}{}{\forall x.(A \vlor B)}{
      \vlin{\fequ}{(x \notin fv(B))}{\forall x.\forall x.(A \vlor B)}{
        \vlin{\wrD}{}{\forall x.(B \vlor (\forall x.A))}{
          \vlhy{\forall x.B}}}}}
  \leadsto
  \vlderivation{
    \vlin{\fequ}{}{\forall x.(A \vlor B)}{
      \vlin{\wrD}{}{\forall x.(B \vlor A)}{
        \vlhy{\forall x.B}}}}
  \end{equation*}

\item $\wrD/\fequ/\acD$:
  \begin{equation*}
  \vlderivation{
    \vlin{\acD}{}{a \vlor B}{
      \vlin{\fequ}{}{(a \vlor a) \vlor B}{
        \vlin{\wrD}{}{(a \vlor B) \vlor a}{
          \vlhy{a \vlor B}}}}}
  \leadsto
  \vlderivation{
    \vlhy{a \vlor B}}
  \end{equation*}

  \begin{equation*}
  \vlderivation{
    \vlin{\acD}{}{a \vlor B}{
      \vlin{\fequ}{}{(a \vlor a) vlor B}{
        \vlin{\wrD}{}{a \vlor (a \vlor B)}{
          \vlhy{a}}}}}
  \leadsto
  \vlderivation{
    \vlin{\wrD}{}{a \vlor B}{
      \vlhy{a}}}
  \end{equation*}

  \begin{equation*}
  \vlderivation{
    \vlin{\acD}{}{\forall x.a}{
      \vlin{\fequ}{(x \notin fv(a))}{\forall x.(a \vlor a)}{
        \vlin{\wrD}{}{(\forall x.a) \vlor a}{
          \vlhy{\forall x.a}}}}}
  \leadsto
  \vlderivation{
    \vlhy{\forall x.a}}
  \end{equation*}

  \begin{equation*}
  \vlderivation{
    \vlin{\acD}{}{\forall x.a}{
      \vlin{\fequ}{(x \notin fv(a))}{\forall x.(a \vlor a)}{
        \vlin{\wrD}{}{a \vlor (\forall x.a)}{
          \vlhy{a}}}}}
  \leadsto     
  \end{equation*}

\item $\wrD/\fequ/\me$:
  \begin{equation*}
  \vlderivation{
    \vlin{\me}{}{(A \vlor B) \vlan (C \vlor D)}{
      \vlin{\fequ}{}{(A \vlan C) \vlor (B \vlan D)}{
        \vlin{\wrD}{}{(C \vlan A) \vlor (B \vlan D)}{
          \vlhy{C \vlan A}}}}}
  \leadsto
  \vlderivation{
    \vlin{\wrD}{}{(A \vlor B) \vlan (C \vlor D)}{
      \vlin{\wrD}{}{(A \vlor B) \vlan C}{
        \vlin{\fequ}{}{A \vlan C}{
          \vlhy{C \vlan A}}}}}
  \end{equation*}

  \begin{equation*}
  \vlderivation{
    \vlin{\me}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
      \vlin{\fequ}{x \notin fv(B \vlan D)}{\forall x.((A \vlan C) \vlor (B \vlan
D))}{
        \vlin{\wrD}{}{(B \vlan D) \vlor (\forall x.(A \vlan C))}{
          \vlhy{B \vlan D}}}}}
  \leadsto
  \end{equation*}

\item $\wrD/\fequ/\mfaD$:
  \begin{equation*}
  \vlderivation{
    \vlin{\mfaD}{}{\forall x.(A \vlor B)}{
      \vlin{\fequ}{}{(\forall x.A) \vlor (\forall x.B)}{
        \vlin{\wrD}{}{(\forall x.B) \vlor (\forall x.A)}{
          \vlhy{\forall x.B}}}}}
  \leadsto
  \vlderivation{
    \vlin{\fequ}{}{\forall x.(A \vlor B)}{
      \vlin{\wrD}{}{\forall x.(B \vlor A)}{
        \vlhy{\forall x.B}}}}
  \end{equation*}

\item $\wrD/\fequ/\mexD$:

  \begin{equation*}
  \vlderivation{
    \vlin{\mexD}{}{\exists x.(A \vlor B)}{
      \vlin{\fequ}{}{(\exists x.A) \vlor (\exists x.B)}{
        \vlin{\wrD}{}{(\exists x.B) \vlor (\exists x.A)}{
          \vlhy{\exists x.B}}}}}
  \leadsto
  \vlderivation{
    \vlin{\fequ}{}{\exists x.(A \vlor B)}{
      \vlin{\wrD}{}{\exists x.(B \vlor A)}{
        \vlhy{\exists x.B}}}}
  \end{equation*}
